*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-05 Code/Manuscript_O2.Rmd
Adedd acknowledgement Sarah Arthur
Changed P .marinus with *Prochlorococcus marinus* when starts sentence.

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Clade LLI includes cultured strain NATL2A, which prefers moderate irradiances typical of between 30 and 100 m depth. Clades LLII and LLIII, including cultured strain SS120, are grouped together as the second oldest phylogenetic lineage diversifying in the *P. marinus* radiation, with an affinity for low light. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, typical of depths from 120 m to 200 m [@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are, as yet, the only cultured strains to have been found in Oxygen Minimum Zones (OMZ).  Some, as yet, uncultured *P. marinus* strains in clades LLV and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021; @lavinNovelLineagesProchlorococcus2010; @goerickeNovelNicheProchlorococcus2000; @garcia-robledoCrypticOxygenCycling2017; @johnsonEnergeticsGrowthKinetics1999]. *P. marinus* LL ecotypes may indeed dominate the phytoplankton within OMZ [@lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @partenskyProchlorococcusAdvantagesLimits2010], where they may be net O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018].

@@

*P. marinus* MED4, clade HLI, growth under 250 µM O~2~ increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), across all spectral wavebands. No growth was observed under any imposed conditions under a 4 h photoperiod. The maximum growth rate (µ~max~) was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 180 µE blue light and 16 h photoperiod.

@@

*P. marinus* SS120 clade LLII/III,  growth under 250 µM O~2~ increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at 16 h red light photoperiod.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white LED, red or blue light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

*P. marinus* MIT9313, clade LLIV, growth under 250 µM O~2~ increased with longer
photoperiods, under low 30 µmol photons m^-2^ s^-1^ peak PAR, (Figure \@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiods. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light at 180 µmol photons m^-2^ s^-1^ peak PAR, under the 8 h photoperiod, consistent with Moore *et al*. [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

# Acknowledgements  {.unnumbered}
Carlie Barnhill (Mount Allison Student) assisted with code for import of multicultivator growth data files.

>>>>

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Clade LLI includes cultured strain NATL2A, which prefers moderate irradiances typical of between 30 and 100 m depth. Clades LLII and LLIII, including cultured strain SS120, are grouped together as the second oldest phylogenetic lineage diversifying in the *P. marinus* radiation, with an affinity for low light. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, typical of depths from 120 m to 200 m [@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are, as yet, the only cultured strains to have been found in Oxygen Minimum Zones (OMZ).  Some, as yet, uncultured *P. marinus* strains in clades LLV and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021; @lavinNovelLineagesProchlorococcus2010; @goerickeNovelNicheProchlorococcus2000; @garcia-robledoCrypticOxygenCycling2017; @johnsonEnergeticsGrowthKinetics1999]. *Prochlorococcus marinus* LL ecotypes may indeed dominate the phytoplankton within OMZ [@lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @partenskyProchlorococcusAdvantagesLimits2010], where they may be net O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018].

@@

*Prochlorococcus marinus* MED4, clade HLI, growth under 250 µM O~2~ increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), across all spectral wavebands. No growth was observed under any imposed conditions under a 4 h photoperiod. The maximum growth rate (µ~max~) was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 180 µE blue light and 16 h photoperiod.

@@

*Prochlorococcus marinus* SS120 clade LLII/III,  growth under 250 µM O~2~ increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at 16 h red light photoperiod.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white LED, red or blue light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

*Prochlorococcus marinus* MIT9313, clade LLIV, growth under 250 µM O~2~ increased with longer
photoperiods, under low 30 µmol photons m^-2^ s^-1^ peak PAR, (Figure \@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiods. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light at 180 µmol photons m^-2^ s^-1^ peak PAR, under the 8 h photoperiod, consistent with Moore *et al*. [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

# Acknowledgements  {.unnumbered}
Carlie Barnhill (Mount Allison Student) assisted with code for import of multicultivator growth data files.
Sarah Arthur (Mount Allison Student) assisted with culturing and setting up Multicultivator runs.
